Case for Support

### 2021-11-16 application/Je-S_sections.tex
Updates from Overleaf

The overall goal of this proposal is to extend, establish, distribute and support Bonsai, a fully integrated software
environment to enable cutting-edge reproducible laboratory systems neuroscience experiments in animal models.

@@

1. Develop, test and deploy a new software pipeline to incorporate external packages into Bonsai's C# environment.

@@

[2825 chars]

@@

@@

>>>>

The overall goal of this proposal is to extend, enhance, maintain and support Bonsai, a fully integrated software
environment to enable cutting-edge reproducible laboratory systems neuroscience experiments in animal models.

@@

1. Develop, test and deploy a new software workflow to incorporate external packages into Bonsai's C# environment.

@@

[2829 chars remaining]

@@

Understanding the brain is a major scientific challenge of our era. To succeed, scientists must be able to explain how the rich repertoire of animal behaviour relates to the activity in neural circuits across different brain regions. This necessitates careful design and control of behavioural experiments wherein experimental animals (e.g. non-human primates, rodents, fish, insects) perform particular tasks while experimenters either record or manipulate neural activity in candidate neural circuits. 
Experiments in behavioural and brain science laboratories require software that integrates and controls hardware from multiple recording devices (video, electrodes for neural activity measurement, sensors), and analysis tools that can interpret large and complex behavioural and neural datasets.

Scientists studying brain and behaviour dedicate the majority of their time to the design of experimental software protocols that control hardware and to data analysis, with least amount spent on data acquisition itself. Moreover, hundreds of neuroscience research groups worldwide develop their own software tools for experimental control and data analysis in different programming languages, leading to inefficiencies resulting from time-loss spent on writing software (often by individuals lacking programming expertise), the re-invention of software with the same functionality and concomitant lack of experimental reproducibility across laboratories.

Here we propose to provide a community software tool that will dramatically increase the efficiency of experimental control and data analysis by providing a new set of functionalities to an existing software platform, Bonsai. Bonsai is an open-source reactive visual programming environment that naturally orchestrates complex networks of sensors and hardware into workflows and allows integration of multiple programming languages into its framework. Thus far, Bonsai has been adopted by hundreds of scientists worldwide to provide interactive experimental control in behavioural and brain sciences. In this proposal, we aim to extend Bonsai’s functionality with a toolbox of online and offline Machine Intelligence tools for analysis of behavioural and neural datasets, and to create an open-access platform for software sharing. Bonsai’s enhanced functionality will enable new types of research via online experimental control, speed up discovery and improve efficiency by (i) providing access to such tools into laboratories lacking expertise, (ii) reducing the need to reinvent the same tools in multiple labs and (iii) standardising the data processing streams, thus increasing reproducibility across laboratories. We believe this effort will enable and accelerate new discoveries in how the brain generates behaviour.

@@

To understand the brain, scientists aim to explain how the rich repertoire of animal behaviour relates to the activity in neural circuits across different brain regions. This necessitates careful design and precise control of behavioural experiments wherein experimental animals perform particular tasks while experimenters either record or manipulate neural activity in candidate neural circuits. Such experiments require software that integrates and controls hardware from multiple recording devices (cameras, electrodes, sensors), and analysis tools that can interpret large and complex behavioural and neural datasets. The lack of standardised tools for experimental control and analysis across the brain-and-behaviour community is incurring major inefficiencies, with disproportionate time allocated to software engineering and writing analysis code, and efforts often replicated across laboratories. Similarly, disparate experimental and analysis protocols makes is almost impossible to replicate experiments between research groups.
Bonsai is an open-source reactive visual programming environment that naturally orchestrates complex networks of sensors and hardware into workflows. Bonsai has been adopted by hundreds of scientists worldwide to provide interactive experimental control in behavioural and brain sciences. Here we aim to extend Bonsai’s functionality with a toolbox of online and offline Machine Intelligence tools for analysis of behavioural and neural data (video-based analysis of behavioural motifs, real-time and off-line analysis of neural signals), and create an open-access platform for software sharing and integration with multiple programming languages. Bonsai’s enhanced functionality will enable new types of research, accelerate discovery and improve efficiency by (i) providing access to such tools for scientists lacking expertise, (ii) reducing duplication of effort, and (iii) standardising the data processing streams, thus increasing reproducibility across laboratories.
[12 chars over]

### 2021-11-16 application/backgroundToTheResource.tex
Updates from Overleaf

Modern systems neuroscience depends on data from carefully designed behavioural experiments in
animal models such as mice frequently combined with invasive electrophysiology and/or
cellular-resolution imaging.  Experimental design might require behavioural feedback (e.g. reward,
lights, sounds, actuated devices\dots) or neural pertubation (e.g. using light-sensitive polarising
molecules) that depends on real-time behaviour or neural state.
%
Furthermore, it has become increasingly apparent that the neural signals that drive complex
behaviour are manifest in a broad population of neurons rather than in single cells.  So current
neural recording emphasises high-dimensional signals from 100s to 1000s of neurons.

This complexity of experiments limits the potential for database-driven science.  The number of
possible experiments is astronomical, and a new question must usually be answered by designing and
conducting a new experiment rather than mining the data from an old one.
%
At present, the field is held back by our ability to efficiently design and analyse new experiments,
and to integrate state-of-the-art data processing and neuroinformatics into experimental designs.
%
\textbf{Our proposal is to enhance, distribute and support a neuroinformatics resource which \dots}
(simplifies data collection and analysis, something about sharing data??), potentially creating the most powerful experimental control and analysis tool available to neuroscientists, psychologists and ethologists worldwide.

@@

\setcounter{figure}{1}

This resource will be built on the platform of Bonsai, an established experimental control software
ecosystem with a long development history and a large user base in systems neuroscience
(Figure~\ref{fig:bonsai}).

A central goal of the software resource proposed here is to address this
problem, by extending Bonsai~(an experimental control software ecosystem;
Figure~\ref{fig:bonsai} and Section~\ref{sec:bonsai}) with state-of-the-art
online (and offline) data analysis methods
(Figure~\ref{fig:proposedBonsaiExtensions} and
Section~\ref{sec:functionality}). \textbf{Bonsai is}
\textbf{unique} among experimental control software in that it allows people
with no programming experience to control sophisticated experiments. Its large
and rapidly growing user base demonstrates the need for this type of software
resources. With the proposed functionality added to Bonsai, its users
will be able to use sophisticated \textbf{machine-learning methods} to control their
experiments and analyse their data. This
\textbf{democratisation} of machine learning
functionality will allow scientists without computational skills to control
sophisticated experiments and analyse their data in unprecedented ways, which
will facilitate \textbf{advances in explaining how the brain gives rise to behaviour}.

Currently Bonsai has its own package manager, and its community of users have
already extended Bonsai's functionality with several contributed packages
(e.g., \ldots \textcolor{red}{Goncalo please reference the best contributed
packages developed by the Bonsai user community}). However, these packages need to
be written in C\#, while most current neuroscience data analysis methods are
written in Python, R or Matlab. We propose to add to Bonsai capabilities to
communicate with software written in these languages
(Section~\ref{sec:softwareInfrastructure}).  These capabilities will allow a
large number of neuroscience data analysis methods to be easily integrated into
the Bonsai ecosystem and will provide Bonsai users a large repertoire of
advanced machine learning methods for their behavioural and neural data
analysis.
%
Using these capabilities we will integrate into Bonsai an initial set of
state-of-the art behavioural and neural data analysis methods
(Section~\ref{sec:functionality}).

Developers of advanced data-analysis methods should become interested in
integrating them into the Bonsai ecosystem, as it will allow their methods to
reach the wide Bonsai user community. In addition this integration will enable
method developers to easily compare the performance of their methods with
that of many methods previously integrated into the Bonsai ecosystem. Method
developers will thus become a new type of Bonsai users, which will contribute
to the \textbf{sustainability} of the proposed resource beyond the period of
BBSRC-BBR funding.

Scientific advances rely on \textbf{reproducibility}, which in turn depends on standardized tools for experimental control and analysis, shared between laboratories (Baker, 2016; Ioannidis, 2005). Although such tools exist in genetics, astronomy, physics and medicine (Fish et al., 2016; Abdalla et al., 2018, CERN Education, Communications and Outreach Group, 2018, Dickinson et al., 2016, Bycroft et al., 2018), they are mostly lacking in neuroscience, which in itself suffers from a paucity of reproducible discoveries (Baker, 2016; Botvinik-Nezer et al., 2020; Button et al., 2013). Bonsai is an excellent tool for reproducible data acquisition and experimental control. For example, an experiment implemented in Bonsai can be replicated in
any laboratory by just sharing a Bonsai configuration file. With the addition of the proposed machine learning functionality Bonsai will extend this
reproducibility to the domain of data analysis
(Section~\ref{sec:reproducibility}).

\begin{figure}
  \includegraphics[width=4in]{figures/proposed_bonsai_extensions.png}
  \caption{Proposed extensions to Bonsai}
  \label{fig:proposedBonsaiExtensions}
\end{figure}

\subsection{Bonsai}
\label{sec:bonsai}

@@

>>>>

To understand the neural basis of behaviour, experiments need to control and quantify animal behaviour whilst recording, and often manipulating, neural activity. 
%
Powerful experimental designs require behavioural feedback (e.g.\ reward, lights, sounds, actuated
devices) or neural perturbation (e.g.\ using light-sensitive polarising
molecules) that depend on real-time behaviour or neural state.
%
Furthermore, it has become increasingly apparent that the neural signals that
drive complex behaviour are manifest in a broad population of neurons rather
than in single cells.  So current neural recording emphasises high-dimensional
signals from 100s to 1000s of neurons monitored using multi-electrode arrays or cellular-resolution imaging, with a wide array of probe configurations and neural targets.
%
This complexity of experiments limits the potential for database-driven progress in systems neuroscience.
%
The number of possible experiments is astronomical, and a new question must
usually be answered by designing and conducting a new experiment rather than
mining the data from an old one.

At present, progress in the field is held back by the challenge of rapidly designing and implementing new experimental protocols, and the difficulty of integrating state-of-the-art data processing and
neuroinformatics into custom experimental designs.
%
\textbf{Reproducibility}, in particular, depends on the availability of standardized tools for experimental control and analysis, shared between laboratories (Baker, 2016; Ioannidis, 2005). Although such tools exist in genetics, astronomy, physics and medicine (Fish et al., 2016; Abdalla et al., 2018, CERN Education, Communications and Outreach Group, 2018, Dickinson et al., 2016, Bycroft et al., 2018), they are mostly lacking in neuroscience, with documented consequences for the reliability of ``discoveries'' (Baker, 2016; Botvinik-Nezer et al., 2020; Button et al., 2013).

@@

@@

Our proposed enhancements will \textbf{enhance, reinforce and support this key neuroinformatics
  resource to streamline reproducible experimental design, data collection and analysis, while making standardised implementations of cutting-edge machine-learning methodologies widely available to the community}, potentially creating the most powerful experimental
control and analysis tool available to neuroscientists, psychologists and
ethologists worldwide.

\input{otherResources.tex}

%

%

% This resource will be built on the platform established by \textbf{Bonsai}.
%

% an established
% experimental control software ecosystem with a long development history and a
% large user base in systems neuroscience (Figure~\ref{fig:bonsai}).

% \subsection{Bonsai}
% \label{sec:bonsai}
%
% We will 
% 
% 
% 
% A central goal of the software resource proposed here is to address this
% problem, by extending Bonsai~(an experimental control software ecosystem;
% Figure~\ref{fig:bonsai} and Section~\ref{sec:bonsai}) with state-of-the-art
% online (and offline) data analysis methods
% (Figure~\ref{fig:proposedBonsaiExtensions} and
% Section~\ref{sec:functionality}).
% Bonsai is unique amongst experimental control systems in that
% it allows users with no programming expertise to design and implement
% sophisticated experiments.
%
% The latest Its large and rapidly growing community of users demonstrates the demand for such software resources.
%

% With the proposed enhancements, users of Bonsai will be able to use sophisticated \textbf{machine-learning methods} to
% control their experiments and analyse their data. This \textbf{democratisation}
% of machine learning functionality will allow scientists without computational
% skills to control sophisticated experiments and analyse their data in
% unprecedented ways, which will facilitate \textbf{advances in explaining how the
%   brain gives rise to behaviour}.

% \subsection{Bonsai}
% \label{sec:bonsai}

### 2021-11-16 application/caseForSupport.tex
Added track record
Updates from Overleaf

\makeatletter
\renewcommand\paragraph{\@startsection{paragraph}{4}{\z@}%
            {-2.5ex\@plus -1ex \@minus -.25ex}%
            {1.25ex \@plus .25ex}%
            {\normalfont\normalsize\bfseries}}
\makeatother
\setcounter{secnumdepth}{4} % how many sectioning levels to assign numbers to
\setcounter{tocdepth}{4}    % how many sectioning levels to show in ToC
\usepackage[small]{titlesec} % format section titles

@@

\maketitle

@@

@@

\section{Details of resource}

>>>>

\makeatletter
\renewcommand\paragraph{\@startsection{paragraph}{4}{\z@}%
            {-2.5ex\@plus -1ex \@minus -.25ex}%
            {1.25ex \@plus .25ex}%
            {\normalfont\normalsize\bfseries}}
\makeatother
\setcounter{secnumdepth}{4} % how many sectioning levels to assign numbers to
\setcounter{tocdepth}{4}    % how many sectioning levels to show in ToC
\usepackage[tiny]{titlesec} % format section titles
\def\headfmt{\color{blue!50!black!50}\bfseries}
\titleformat*{\section}{\headfmt}
\titlespacing*{\section}{0pt}{*1}{*0}[-4em]
\titleformat*{\subsection}{\headfmt}
\titlespacing*{\subsection}{0pt}{*0.5}{*0}
\titleformat*{\subsubsection}{\headfmt}
\titlespacing*{\subsubsection}{0pt}{*0}{*0}
\titleformat*{\paragraph}{\headfmt}
\titlespacing*{\paragraph}{0pt}{*0.5}{*0}

@@

% \maketitle

@@

\section*{\large Case for support}
\vskip 0.5ex
We propose to extend, enhance, maintain and support \textbf{Bonsai}, a fully
integrated software environment to \textbf{enable cutting-edge reproducible
  laboratory systems neuroscience experiments in animal models}, with a
particular emphasis on \textbf{machine-intelligence-enabled, real-time
  neuroinformatics} methods.

@@
